### 2023-05-24 manuscript/main.qmd
discussion part 2 and proofreading checks

---
title: "Bayesian multi-proxy reconstruction of early Eocene latitudinal temperature gradients"
author:
  - "---"
  - Kilian Eichenseer^1^ and Lewis A. Jones^2^
  - "---"
  - "^1^Department of Earth Sciences, Durham University, South Road, DH1 3LE, Durham, United Kingdom"
  - "^2^Centro de Investigación Mariña, Grupo de Ecoloxía Animal, Departamento de Ecoloxía e Bioloxía Animal, Universidade de Vigo, 36310 Vigo, Spain."
  - "---"
  - "**Corresponding author:** kilian.eichenseer@durham.ac.uk"
abstract: \newpage Abstract
format:
  docx:
    reference-doc: template.docx
    highlight-style: github
    fig-dpi: 900
  pdf:
    documentclass: scrartcl
    papersize: letter
    geometry:
      - top=25.4mm
      - bottom=25.4mm
      - left=25.4mm
      - right=25.4mm
      - heightrounded
header-includes:
  - \usepackage{newunicodechar}
  - \newcommand{\$^\circ$C}{\textdegree C}
bibliography: references.bib
csl: climate-of-the-past.csl
editor: source
---

```{=html}
<!-- discarded titles:
Bayesian multi-proxy reconstruction of latitudinal temperature gradients applied to the Early Eocene
A Bayesian model of latitudinal temperature gradients applied to the Early Eocene
Bayesian multi-proxy reconstruction of latitudinal temperature gradients for the Early Eocene
Bayesian multi-proxy reconstruction of the early Eocene latitudinal temperature gradient
Combining geochemical and ecological proxies...
Modelling latitudinal temperature gradients from geological proxies: An early Eocene example
"Improving palaeoclimatic reconstructions with ecological data: How hot was the early Eocene?"
"Integrating ecological and geochemical proxies for palaeoclimatic reconstruction: An early Eocene example"
"The Early Eocene Climate Optimum reconstructed with Bayesian multi-proxy integration"
"An Early Eocene temperature gradient modelled with geological and ecological proxies"
-->
```
Accurately reconstructing large-scale palaeoclimate patterns from sparse local records is critical for understanding the evolution of Earth's climate. Challenges arise particularly from the patchiness and uneven sampling of palaeoclimate data, and from the disparate nature of the proxy data. Geochemical data usually provide temperature estimates via transfer functions informed by experiments, and transfer functions based on the climatic requirements of modern taxa exist for some fossil plant assemblages. In contrast, most ecological and lithological data, e.g. coral reefs and evaporites, only convey information on broad climatic requirements. Historically, most large-scale proxy-based reconstructions have been based on either geochemical or ecological data, but few approaches exist for combining multiple proxy types into a single, quantitative reconstruction. Large spatial gaps in existing proxy records have often been bridged by simple averaging, without taking into consideration the spatial distribution of the samples, leading to biased temperature reconstructions. Here, we present a Bayesian hierarchical model to integrate ecological data with established geochemical proxies into a unified quantitative framework, bridging gaps in the latitudinal coverage of proxy data. We apply this approach to the early Eocene climatic optimum (EECO), the interval with the warmest sustained temperatures of the Cenozoic. Assuming the conservation of thermal tolerances of modern coral reefs and mangrove taxa, we establish broad sea surface temperature ranges for EECO coral reef and mangrove sites. We integrate these temperature estimates with the EECO geochemical shallow marine proxy record to model the latitudinal sea surface temperature gradient and global average temperatures. Our results confirm the presence of a flattened latitudinal temperature gradient and unusually high polar temperatures during the EECO, which is supported by high-latitude ecological data. Integrating multiple types of proxy data and adequate prior information has the potential to substantially reduce uncertainty in palaeoclimate reconstructions, allowing for unbiased temperature estimates from sparse data.

# Keywords

Palaeoclimate, latitudinal temperature gradients, temperature proxies, Eocene, spatial bias, Bayesian

<!--Geochemical data such as oxygen isotopes (d18O), tetraether indices (TEX86) and magnesium calcium ratios (Mg/Ca) are routinely used to reconstruct ocean temperatures in deep time. However, limited spatial coverage and disagreements between geochemical proxies compromise the accuracy of these reconstructions. Geological data such as coral reefs, mangroves, and evaporites have the potential to supplement geochemical data by improving spatial coverage and constraining temperature estimates. Historically, these data have been used to qualitatively inform upon palaeoclimatic conditions at broad spatial scales, yet no transfer functions exist to convert geological data into temperature estimates. Quantitative temperature reconstructions have therefore not made use of these data. Likewise, the large spatial gaps in existing proxy records have often been bridged by simple averaging, without taking into consideration the spatial distribution of the samples. Here, we present a Bayesian hierarchical model to integrate geological data–with established temperature proxies–into a unified quantitative framework. We apply this approach to the early Eocene climatic optimum (EECO), the interval with the warmest sustained temperatures of the Cenozoic. Assuming the conservation of thermal tolerances of modern coral reefs and mangrove taxa, we establish broad sea surface temperature ranges for EECO coral reef and mangrove sites. We integrate these temperature estimates with the EECO geochemical shallow marine proxy record to model the latitudinal temperature gradient and its uncertainty. Our results confirm the presence of a flattened latitudinal temperature gradient and unusually high polar temperatures during the EECO. We show that the inclusion of ecological data can substantially reduce the uncertainty on temperature estimates in climate zones lacking geochemical data. -->

# Introduction

Understanding the long-term evolution of Earth's climate system and contextualising current global warming relies on accurate reconstructions of past climates [@Royer2004; @Burke2018]. Recent advances in the synthesis of palaeoclimate data [e.g. @Veizer2015; @Hollis2019; @song2019; @Grossman2022; @Judd2022] are offering unprecedented insights into the complex and dynamic nature of the Earth's climate system, but a fundamental challenge remains: the proxy record of past climates is spatially incomplete and afflicted by imperfect preservation and uneven sampling [@judd2020; @Jones2022; @Judd2022].

Whilst geochemical proxy data can provide robust estimates of palaeotemperature at local scales, recent work has demonstrated that spatial biases in the geochemical proxy record can lead to spurious estimates of regional (e.g. latitudinal temperature gradients) and global temperatures   [@judd2020; @Jones2022]. At the most basic level, this can be driven by two factors: (1) missing data for some regions (e.g. no high-latitude data); or (2) overrepresentation of other regions (e.g. a high proportion of samples from tropical areas). The latter can be addressed through the down-sampling of data or restricting analyses to specific regions (e.g. @song2019). However, in order to robustly infer regional or global-scale patterns from an incomplete record, spatial gaps must ultimately be bridged. One common approach, which requires no additional computation, is the spatial visualisation of proxy-derived temperatures against latitude, showing broad latitudinal temperature trends [@Hollis2019; @Vickers2021]. Interpolation is also sometimes used to bridge spatial gaps in palaeoclimate data [@Taylor2004], taking advantage of the autoregressive nature of climatic data: much of the information on the climate of any given location is contained in the climate data of nearby locations [@Reynolds1994]. Adding to this, some proxy-based reconstructions use statistical modeling to infer palaeoclimatic patterns. For example, polynomial regression [@Bijl2009] and cosine functions [@Inglis2020] have been used to reconstruct latitudinal temperature gradients, and 2D-reconstructions of surface temperatures have been created with Gaussian process regression [@Inglis2020]. These approaches work well for interpolating relatively densely-sampled data, but the absence of constraints on the modeled parameters means that such models can produce unrealistic temperature estimates when extrapolating from sparse data. Statistical modeling in the Bayesian framework can help overcome this problem by requiring the explicit specification of priors for the model parameters, which can be used to express physical constraints [@Chandra2021].

Spatial gaps in the palaeoclimate record can also be addressed through the integration of additional data. For example, lithological and fossil data can be used to infer past climatic conditions based on analogous modern sediments [@Chandra2021], or based on the premise that the climatic requirements of ancient taxa, biological traits, or ecological communities were similar to those of their nearest modern relatives [@Peppe2011; @Royer2012; @Salonen2019]. Despite this potential, the integration of geochemical proxy data with other sources of information (e.g. ecological data) has rarely been realised in a rigorous, quantitative framework [@Burgener2023].

Here, we present a novel Bayesian hierarchical model that combines quantitative proxies and ecological constraints into a fully quantitative model of the latitudinal gradient of sea surface temperatures, bridging spatial gaps in sparsely sampled climate data. This model expands upon existing, spatially explicit palaeoclimate reconstructions  by allowing for the integration of (1) prior information based on physical principles and on the observed, modern sea surface temperature distribution, and of (2) geochemical and ecological climate proxies in a common, quantitative framework. We use a generalised logistic function to accurately infer the shape of the temperature gradient despite a patchy latitudinal coverage, and test the robustness of this method using downsampled, simulated temperature gradients.

We apply this model to the record of the early Eocene climatic optimum (EECO), combining a compilation of geochemical proxies [@Hollis2019], mangrove communities [@Popescu2021], and coral reefs [@Zamagni2012], using a nearest-living-relative approach [see e.g. @Greenwood2017] to establish broad temperature ranges for the ecological data. We chose the EECO to demonstrate the application of the model due to its significance as the interval with the warmest sustained temperatures of the Cenozoic [@Pross2012], rendering it a potential analogue for extreme climate warming scenarios [@Burke2018]. Our integrative approach allows us to shed new light on the long-standing dispute on the steepness of the early Eocene temperature gradient [@Sloan1990; @Markwick1994; @Huber2011; @Tierney2017; @Inglis2020].

::: {#table-1}
![](../figures/Table1.png){width="16.25cm"}
Table 1: Inferred latitudinal sea surface temperature (SST) gradients for the early Eocene (EE) or the EECO, as shown in earlier, proxy-based studies. For comparison, a gradient derived from a climate model ensemble is also shown.

:::

# Materials & Methods

```{r methods block - for in-line code, echo = FALSE}
setwd('..') # move up to parent folder of /manuscript

# data
obsmat <- readRDS("data/processed/observation_matrix_for_EECO_model.rds")
n_geochem <- nrow(obsmat)
n_geochem_location <- length(unique(obsmat$sample))

```
## Modern sea surface temperatures

## Geochemical data

Geochemical climate proxy data were extracted from a latest Paleocene and early Eocene compilation [@Hollis2019]. This compilation provides data on four different geochemical proxies for reconstructing seawater temperature: &delta;^18^O, &Delta;47, Mg/Ca and $TEX_{86}$. For our analyses, this dataset was restricted to the EECO (defined as 49.1 - 53.8 Ma) and samples from the continental shelf. Recrystallised &delta;^18^O samples were also excluded as secondary diagenetic calcite precipitated after deposition can bias isotope measurements and offset temperature values [@Schrag1999]. This filtering resulted in most &delta;^18^O samples being excluded from the dataset (retaining 8 out of 152). After data filtering, `r n_geochem` geochemical proxy samples from `r n_geochem_location` locations remained. For a detailed description of each proxy see @Hollis2019.

## Ecological data

**Coral reefs.** Today, shallow warm-water coral reefs are limited to tropical and subtropical latitudes (\~34$^\circ$N--32$^\circ$S), with minimum sea surface temperature tolerances (\~18$^\circ$C) being the primary constraint on this distribution [@johannes1983; @kleypas1999; @yamano2001]. As coral reefs reside at the upper thermal limit of the oceans today, their maximum sea surface temperature tolerance is less well-constrained, with some studies suggesting up to 35.6$^\circ$C in the geological past [@jones2022reef]. Nevertheless, coral reefs have frequently been recognised as tracers of past (sub-)tropical conditions [@ziegler1984; @kiessling2001]. During the Eocene, coral communities and reefs expanded across tropical and temperate latitudes, with some communities found up to palaeolatitudes of 43 $^\circ$N [@Zamagni2012]. Using a compilation of Paleocene--early Eocene coral reefs and community localities [@Zamagni2012], we generated quantitative sea surface temperature estimates for the ECCO. To do so, we extracted localities from the compilation that are inferred to be Ilerdian (early Eocene) coral reefs, and that could be confidently assigned to the EECO. We excluded coral knobs and coral-bearing mounds which might have broader climatic limits than coral reef ecosystems. This filtering resulted in four unique coral reef localities remaining for the EECO, all of which conform to the modern latitudinal range of coral reefs (\<34$^\circ$ N). Subsequently, we used statistically derived temperature limits (minimum = 21$^\circ$C, average = 27.6$^\circ$C, maximum = 29.5$^\circ$C) from the published literature [@kleypas1999] to define a normal probability distribution of potential temperature values for coral reef localities. This normal probability distribution was defined with a mean of 27.6 and a standard deviation of 2.125, placing 97.5% of the probability density above the minimum. As the distribution of modern corals is skewed towards warmer temperatures, this approach results in 16.5% of the probability being placed on temperatures \> 29.5$^\circ$C, allowing for the possibility that Eocene coral reefs were adapted to warmer conditions than present-day coral reefs.

**Mangroves.** Mangroves are distributed throughout the tropics and subtropics today. While factors beside sea surface temperatures (SST) influence the distribution of mangroves, empirical, lower temperature limits have been established for the genera *Avicennia* (15.6$^\circ$C) and *Rhizophora* (20.7$^\circ$C) [@Quisthoudt2012]. Both *Avicennia* and members of the Rhizophoraceae family were widespread in the early Eocene, but only *Avicennia* occurred at polar latitudes [@Suan2017; @Popescu2021]. Assuming that Eocene members of these mangrove taxa conform to similar climatic requirements as their modern relatives, the presence and absence of *Avicennia* and Rhizophoraceae pollen can be used as a palaeotemperature indicator. For this analysis, published mangrove occurrence data were taken from @Popescu2021, and converted to quantitative temperature estimates. From this data, we identify two types of pollen assemblages which we ascribe different temperature distributions:

1)  *Avicennia*-only assemblages ($n=2$): the absence of Rhizophoraceae is indicative of temperatures being between 15.6$^\circ$C (lower temperature limt of *Avicennia*) and 20.7 (lower temperature limit of *Rhizophora*). However, a value of 22.5$^\circ$C is ascribed as the upper temperature limit here as *Rhizophora* is rare below this temperature. We define the *Avicennia*-only temperature distribution as a normal distribution with a mean of 19.05 and a standard deviation of 1.725, resulting in 95% of the probability density being placed within the temperature limits.

2)  *Avicennia* and Rhizophoraceae assemblages ($n=5$): the presence of both groups suggests that the locality should have a minimum temperature of 20.7$^\circ$C (lower temperature limit of *Rhizophora*). As the upper thermal limits of *Aviciennia* and *Rhizophora* are not well established in @Quisthoudt2012, we assign the same maximum temperature limits (29.5$^\circ$C) as coral reef localities, because mangroves are also widely distributed throughout tropical regions. Consequently, we define the temperature distribution for this locality as a normal distribution with a mean of 25.1 and a standard deviation of 2.2, with 95% probability density within the temperature limits.

::: {#fig-1}
![](../figures/fig_1.png){width="15cm"}

Palaeogeographic distribution of the geochemical and ecological data compilation used in this study. <!--  I cut repetition from the methods, but maybe cite where the map is coming from. 
Palaeocoordinates of samples were estimated via the palaeoverse R package [version 1.2.0,@jones2023]. To do so, the @merdith2021 Global Plate Model was used to palaeorotate present-day coordinates to 51.2 Ma. --> Map is presented in the Robinson projection (ESRI:54030).
:::

## Palaeogeographic reconstruction

The palaeogeographic distribution of geochemical and ecological data was reconstructed using the @merdith2021 plate rotation model via the palaeoverse R package [version 1.2.0, @jones2023]. The midpoint age of the EECO (51.2 Ma), along with the present-day coordinates of geochemical and ecological data, were used for palaeogeographic reconstruction.

## Bayesian framework

**Model structure (Fig 2).** We model the mean temperature ($\mu$) at location $j$ as a function of absolute latitude ($abs(l)$) with a logistic regression (also known as "growth curve" or "Richard's curve") of the form:

$$\mu_{j}  \sim N(\nu_j, \sigma), $${#eq-mu}

$$\nu_{j} = A~+\frac{K-A}{e^{B(abs(l_j)-M)}}~, ~~~~~ j = 1,...,n,  $${#eq-nu} where $A$ and $K$ denote the lower and upper asymptote, respectively, $M$ specifies the latitude of maximal growth, i.e. the latitude around which temperature falls most steeply with latitude, $B$ denotes the growth rate (Fig. 2), $\sigma$ denotes the residual standard deviation, and $n$ denotes the number of locations.

We infer $\mu_j$ from $m$ individual temperature observations $t_{i = 1,...m}$, derived from geochemical data, at location $j$ as

$$t_{i,j}  \sim N(\mu_j, \sigma_j), ~~~~~ i = 1,...,m,  $${#eq-obs} where $m$ is the number of observations at each location, and $\sigma_j$ is the estimated standard deviation of the temperatures at location $j$.

Similarly, $\mu_j$ is inferred for locations with ecological proxies from the associated normal temperature distributions with a given mean and standard deviation, $t_{\mu,j}$ and $t_{\sigma,j}$, as

$$t_{\mu,j}  \sim N(\mu_j,t_{\sigma,j}).  $${#eq-dist}

This structure implies that $\mu_j$ is not fixed at the mean proxy temperature at location $j$, <!--of the temperatures $t_{i=1,...,m,~j}$ of location $j$--> but is drawn towards the overall logistic regression curve, i.e. towards $\nu_j$. The pull towards $\nu_j$ tends to be high when $m$ is low, when the observations $t_{i=1,...,m,~j}$ are scattered, i.e. $\sigma_j$ is high, and/or when the overall standard deviation $\sigma$ is low. In practice, this has the desirable consequence that locations with few observations and large temperature differences between observations have less influence on the overall regression than well-sampled locations with consistent temperature observations.

**Priors.** In the Bayesian framework, priors need to be placed on the unknown parameters of a model. We placed weakly informative, conjugate inverse-gamma priors on $\sigma$ and $\sigma_{j=1,..n}$: $$\sigma  \sim \sqrt{Inv-Gamma(\alpha+\frac{n}{2},\beta + 0.5\times(\mu_j-\nu_j))}, ~~~~~ j = 1,...,n,  $${#eq-sigma}
$$\sigma_j  \sim \sqrt{Inv-Gamma(\alpha+\frac{m}{2},\beta + 0.5\times(t_{i,j}-\mu_j))}, ~~~~~ i = 1,...,m,~~~~~j = 1,...,n. $${#eq-sigmaj}
We set $\alpha = \beta = 1$, allowing these priors to be quickly overwhelmed by the data as $n$ and $m$ increase, as we have little *a priori* knowledge of these parameters.

In contrast, we put informative priors on the regression coefficients $A$, $K$, $M$ and $B$, based on physical principles, and vaguely based on the modern climate system:

**A.** Predicted seawater surface temperatures are not allowed to be $<< -2^\circ C$, the freezing point of sea water. The highest prior density of $A$ is placed around $0^\circ C$, and it slowly tapers off towards higher temperatures. This shape is achieved by placing a skew-normal prior on the lower asymptote, specified as $$A \sim SN(\xi=-3.0,\omega=12,\alpha_{SN}=30),$${#eq-A}
where $\xi$, $\omega$, and $\alpha_{SN}$ are the location, scale and shape parameters.

**K.** Input of solar energy decreases from the tropics to the poles. Hence, the latitudinal temperature gradient is broadly negative, i.e. temperature decreases with absolute latitude. This is achieved by setting $K \ge A$. The prior on the upper asymptote $K$ is a truncated normal distribution with the mean set to $K$ of the modern SST gradient, with a broad standard deviation: $$K \sim TN(\mu_{TN} = 28, \sigma_{TN} = 10, \alpha_{TN} = A, \beta_{TN}= \infty)$${#eq-K}
The distribution is truncated to the left at $\alpha_{TN}=A$, but not truncated to the right ($\beta_{TN}$).

**M.** The steepness of the gradient is presumed to be highest in mid-latitudes; this is expressed with a normal prior on $M$ with the mean set to $M$ of the modern SST gradient: $$M \sim N(42,10)$${#eq-M}

**B.** The steepness or growth rate $B$ of the gradient is constrained to be $\ge0$ and to not be exceedingly high, as oceanic and atmospheric heat transfer is bound to prevent very abrupt SST changes across latitudes. A gamma-distributed prior of the form $$B \sim Gamma(\alpha_G=4.3,\beta_G=30)$${#eq-B}
was placed on $B$. The shape and rate parameters $\alpha_G$ and $\beta_G$ were chosen such that the highest prior density is at $B$ of the modern SST gradient, $0.11$.

## Model validation

To test whether our logistic regression model can adequately describe different latitudinal temperature gradients at various sample sizes, we generated four idealised gradients that emulate potential climatic states throughout Earth's geological history: extreme icehouse, icehouse, greenhouse and extreme greenhouse [@Frakes1992]. We then randomly sampled (1000 iterations) these gradients using increasing sample sizes (5, 10, and 20) and reconstructed the latitudinal temperature gradient using our model for each of these sample sizes and gradient types. Using the same idealised gradients, we also tested whether our model could accurately reconstruct latitudinal temperature gradients using the palaeogeographic distribution of Eocene samples (*n* = 34), providing an empirical, examplary distribution that captures both limited sample size and skewed geographic origins of samples. To assess how well the modeled gradients, reconstructed from limited sampling, fit the idealised gradients, we calculated an $R^2$ for Bayesian regression models [@Gelman2019]. For every iteration from the posterior, we intercepted the modeled and the idealised gradient in intervals of 1$^\circ$ latitude, and calculated the $R^2$ based on these values. We report the median, and $2.5^{th}$ and $97.5^{th}$ percentile of the resulting $R^2$ measures.

To test whether our model can accurately depict the shape of the modern sea surface temperature gradient, and to facilitate comparison with the Eocene gradient, we applied our model to annual sea surface mean temperatures from Bio-Oracle [@Assis2018], down-sampled to a $1^\circ \times 1 ^\circ$ raster (*n* = 46,131). The $R^2$ for the modern gradient was calculated as above [@Gelman2019], comparing the modeled gradient and the empirical temperature averages in 1$^\circ$ latitude bins. Only the medians are reported for the modern gradient, as the 95% credible intervals are extremely narrow due to the high precision of the posterior estimates.

To reconstruct the idealised gradients and the modern gradient, we used a simplified, non-hierarchical version of our model, as every location is associated with only one temperature value, making the hierarchical structure superfluous. To achieve this, we substituted temperature ($t_j$) for $\mu_j$ in [@eq-mu] and [@eq-sigma].

## Parameter estimation

We estimated the posterior distributions of the model parameters using a Markov chain Monte Carlo (MCMC) algorithm, written in R. Specifically, we sampled the unknown parameters $A$, $K$, $M$ and $B$ with Metropolis-Hastings, and used Gibbs sampling to estimate all other unknown parameters [see @Gilks1995; @Gelman2013]. Posterior inference on the modern gradient is based on four chains with 60,000 iterations each, 10,000 of which were discarded as burn-in. Every 10th iteration was retained, resulting in a total of 20,000 iterations with low autocorrelation. The re-sampled, simulated gradients were modeled in one chain with 10,000 iterations for each of the 1,000 random samples. 5,000 iterations each were discarded as burn-in, and every 25th iteration was kept, resulting in a total of 200,000 iterations across all 1,000 model runs. For the simulated gradients with an Eocene sampling distribution, a single chain with 250,000 iterations was used, thinned to 10,000 iterations after burn-in. For the Eocene model, we ran four chains with 600,000 iterations each, discarding 100,000 as burn-in and keeping every 100th iteration, as the hierarchical model structure results in higher autocorrelation of the chains. The Eocene posterior inference is thus based on a total of 20,000 iterations with low autocorrelation (effective multivariate sample size for $A$, $K$, $M$ and $B$ is \> 18,000). Trace plots of the MCMC chains indicate convergence and good mixing of the chains (Supplementary Figure S1).

## Processing of model results

```{=html}
<!--
- Calculate lat gradient
- Calculate global T
[Fig. 4]
-->
```
Modeled sea surface temperature estimates were generated with [@eq-nu], calculating the sea surface temperatures at any latitude with the parameter estimates of each iteration from the posterior. The median, $2.5^{th}$ and $97.5^{th}$ percentiles of temperatures where then taken from all temperature estimates obtained at the latitudes of interest.

The latitudinal gradient is calculated as the difference between the modeled temperature at the equator (0$^\circ$ latitude) and at the poles (90$^\circ$ absolute latitude). To facilitate comparison with earlier estimates, we also calculate the gradient with the temperature at the polar circle (66.6$^\circ$ absolute latitude) being used instead of at the poles. Given the sigmoidal shape of the modern as well as the Eocene gradient (see Fig. 4), these results are broadly comparable to a gradient inferred from the zonal average of equatorial and high-latitude temperatures, as is done in some earlier studies [@Evans2018].

Differences between Eocene and modern temperatures at a certain latitude were calculated by randomly pairing all iterations of the posterior from the Eocene and modern temperature gradient model, calculating the Eocene and modern temperature using the respective iterations, taking the difference, and then calculating the median ($2.5^{th}$, $97.5^{th}$ percentile) from all pairs of iterations.

Global average temperatures with 95% credible intervals were calculated by taking the weighted mean of the median ($2.5^{th}$, $97.5^{th}$ percentile) temperature estimates in 1$^\circ$ latitudinal bins. The weights were set to the proportion of global surface area in each latitudinal bin, i.e. decreasing with increasing latitude as: $$weights = sin(\alpha_{1,i}) - sin(\alpha_{2,i}) ,$${#eq-weights} where $\alpha_1$ is the upper, and $\alpha_2$ is the lower latitudinal boundary of bin $i$.

# Results

```{r results block - for in-line code, echo = FALSE}
setwd('..') # move up to parent folder of /manuscript

# temperature gradients
tdiff_lat0 <- readRDS("results/eeco/eocene_modern_difference_equator.rds")
tdiff_lat90 <- readRDS("results/eeco/eocene_modern_difference_poles.rds")
eocene_gradient <- readRDS("results/eeco/eocene_overall_gradient.rds")
eocene_proxy_gradient <- readRDS("results/eeco/eocene_proxy_gradient.rds")
eocene_gradient_polar_circle <- readRDS("./results/eeco/eocene_overall_gradient_polar_circle.rds")

modern_gradient <- readRDS("results/modern/modern_overall_gradient.rds")
modern_eocene_gradient <- readRDS("results/modern/modern_eocene_sampling_overall_gradient.rds")
modern_eocene_gradient_diff <- readRDS("results/modern/modern_eocene_sampling_overall_gradient_diff.rds")
proxy_uncertain_diff_percent <- readRDS("results/eeco/proxy_vs_full_uncertain_diff_percent.rds")

eocene_north_south_gradient_diff <- readRDS("./results/SM/eocene_north_south_gradient_diff.rds")

modern_gradient_empirical <- readRDS("results/modern/modern_empirical_gradient.rds")

# sdy
eeco_sdy <- readRDS("results/eeco/eeco_sdy.rds")
modern_eeco_sample_sdy <- readRDS("results/modern/modern_eeco_sample_sdy.rds")
modern_sdy <- readRDS("results/modern/modern_sdy.rds")

# modern R2
R2_modern <- readRDS("results/modern/modern_model_empirical_R2_gelman.rds")

# global average temperatures
motemp_global <- readRDS("results/modern/global_mean_modern.rds")
eotemp_global <- readRDS("results/eeco/global_mean_eeco.rds")
moeotemp_global <- readRDS("results/eeco/global_mean_modern_eocene_sampling_eeco.rds")
moeotemp_global_diff <- readRDS("results/eeco/global_mean_modern_eocene_sampling_difference.rds")
eoceneproxy_global_diff <- readRDS("results/eeco/global_mean_eocene_eoceneproxy_difference.rds")
eeco_north_global <- readRDS("results/SM/global_mean_northern_eeco.rds")
eeco_south_global <- readRDS("results/SM/global_mean_southern_eeco.rds")
eeco_south_north_global_diff <- readRDS("results/SM/global_mean_diff_south_north_eeco.rds")

```

## Model validation and modern gradient

::: {#fig-2}
![](../figures/fig_2.png){width="15cm"}

Model reconstructions of simulated latitudinal temperature gradients at various sample sizes. Each column depicts a different reconstruction for given sample sizes: 5, 10, 20, and 34 (latitudes of EECO samples). Each row depicts a different simulated latitudinal temperature gradient that represents idealised climatic states: extreme icehouse, icehouse, greenhouse, extreme greenhouse. The black line illustrates the simulated gradient. The blue line depicts the reconstructed gradient represented by the median sea surface temperature value estimated from 1,000 model runs with different random samples (first three columns), and a single run with the EECO latitudinal sampling distribution (fourth column). The blue shadings depict the 90%, 95%, and 99% credible intervals. Bold black text within each panel depicts the coefficient of determination ($R^2$) for estimating goodness of fit between the simulated and modeled gradient. The median (50%) $R^2$ value along with the 2.5% and 97.5% confidence intervals from the 1,000 model runs are shown. Each gradient is depicted in absolute latitude.
:::

```{=html}
<!-- - Show that the 4-parameter logistic function can adequatly describe the modern temperature gradient
- Show that the modern gradient can be reproduced with Eocene sampling distribution
-->
```

::: {#fig-3}
![](../figures/fig_3.png){width="10cm"}

Present-day latitudinal temperature gradient. The present-day empirical latitudinal temperature gradient (mean sea surface temperature) is depicted as a black line, and the gradient estimated by the Bayesian model is shown in turquoise. Grey points depict the individual cell values of the Bio-ORACLE grid of mean sea surface temperatures, which were used to infer the empirical and the modeled gradient.
:::

Our Bayesian model is able to accurately model a range of idealised temperature gradients, ranging from extreme icehouse to "super greenhouse" scenarios (Fig. 2). Random latitudinal sampling results in highly accurate reconstructions already at a sample size of 10 for the icehouse scenarios (95 % CI of $R^2$ > 0.9). Greenhouse scenarios require more random samples to accurately predict high-latitude temperatures. This is because in the absence of high-latitude samples, the modeled gradient is heavily influenced by the priors, which we based on the modern, the only empirically known latitudinal temperature gradient. A sampling distribution resembling that of the early Eocene data set used in this study allows for a highly accurate reconstruction of even the extreme greenhouse scenario (95 % CI of $R^2$ > 0.95).

The average, modern temperature gradient can be closely approximated with our model when using the full modern SST dataset (Fig. 3); almost all of the variation in the empirical median temperatures in bins of 1$^\circ$ absolute latitude is explained by the modeled gradient ($R^2 = `r R2_modern`$). The empirical gradient spans `r modern_gradient_empirical`$^\circ$C from the equator to the poles, the modeled gradient is only very slightly higher at `r modern_gradient$"50%"`$^\circ$C. The modern, global men sea surface temperature based on our modeled gradient is `r motemp_global$"50%"`$^\circ$C.

<!-- REMOVED
Reducing the amount and the spatial coverage of data points to match the early Eocene sampling distribution ($n = 34$) still results in a close match of the modeled gradient with the empirical gradient (Fig. 3b). The 95% credible interval (CI) of reconstructed gradients (Fig. 3b) is noticeably narrower than the spread of empirical temperature values (Fig. 3a). The modern, latitudinal gradient, modeled with all data, spans `r modern_gradient$"50%"`$^\circ$C from the equator to the poles. When modeled with only the early Eocene sampling locations, the median temperature gradient is `r modern_eocene_gradient$"50%"` (95% CI: `r modern_eocene_gradient$"2.5%"` - `r modern_eocene_gradient$"97.5%"`)$^\circ$C. A random sampling of modern temperatures at latitudes corresponding to the sampled Eocene palaeolatitudes would thus allow the reconstruction of the modern latitudinal temperature gradient with good accuracy, with expected deviations from the actual gradient as low as `r modern_eocene_gradient_diff$"2.5%"` to `r modern_eocene_gradient_diff$"97.5%"`$^\circ$C (95 % CI).

Additional model runs with simulated SST data show that latitudinal gradients following a flat, linear or quadratic function can also be adequately reconstructed by our model (Fig. S4). The generalised logistic function underlying the reconstruction leads to the inferred gradient being pushed towards a sigmoidal shape in the absence of proxies (see very high latitudes in Fig. S4a,c). This is intended, as in the absence of data, the shape of the reconstructed gradient should be broadly similar to the sigmoidal shape of the modern SST gradient, as it is the only empirically observed SST gradient.

Modeling the latitudinal temperature gradient facilitates accurate reconstruction of global mean surface temperatures (GMST). The modern, global mean sea surface temperature (GMsST) estimate is `r motemp_global$"50%"`$^\circ$C when using the full, modern data set. When reduced to the early Eocene sampling distribution, the modern GMsST is estimated at `r moeotemp_global$"50%"` (`r moeotemp_global$"2.5%"` - `r moeotemp_global$"97.5%"`)$^\circ$C. The deviations from the full modern estimate are again very low, with a 95 % credible interval of `r moeotemp_global_diff$"2.5%"` to `r moeotemp_global_diff$"97.5%"`$^\circ$C.
-->

## EECO reconstruction

The modeled Eocene temperature gradient is starkly different from the modern (Fig 4). Modelled, median equatorial temperatures are `r tdiff_lat0$"50%"` (`r tdiff_lat0$"2.5%"` - `r tdiff_lat0$"97.5%"`)$^\circ$C higher for the EECO, and polar temperatures are higher by `r tdiff_lat90$"50%"` (`r tdiff_lat90$"2.5%"` - `r tdiff_lat90$"97.5%"`)$^\circ$C. This results in a flattened latitudinal temperature gradient of `r eocene_gradient$"50%"` (`r eocene_gradient$"2.5%"` - `r eocene_gradient$"97.5%"`)$^\circ$C for the EECO, as opposed to `r modern_gradient$"50%"`$^\circ$C for the modern. To facilitate the comparison with latitudinal gradients reported in the literature, which sometimes do not report temperatures in very high latitudes, we report also the EECO gradient between the equator and the modern-day polar circle (66.6$^\circ$), which is slightly lower at `r eocene_gradient_polar_circle$"50%"` (`r eocene_gradient_polar_circle$"2.5%"` - `r eocene_gradient_polar_circle$"97.5%"`)$^\circ$C.

The high variability of EECO palaeotemperature proxies, particularly in the mid-latitudes, and the scarcity of high-latitude data, result in substantial uncertainties in the modeled temperature gradient. This is reflected in the residual standard deviation ($\sigma$) of the EECO gradient, `r eeco_sdy$"50%"` (`r eeco_sdy$"2.5%"` - `r eeco_sdy$"97.5%"`)$^\circ$C, more than twice as high than $\sigma$ of the modern gradient, `r modern_sdy$"50%"`. This signifies that the early Eocene data fit less well to the logistic, latitudinal gradient model, which can also be seen from the drastic departure of some of the proxy data from the gradient estimates (Fig. 4).

The early Eocene GMsST is estimated at `r eotemp_global$"50%"` (`r eotemp_global$"2.5%"` - `r eotemp_global$"97.5%"`)$^\circ$C, `r  round(as.numeric(eotemp_global$"50%") - as.numeric(motemp_global$"50%"),1)`$^\circ$C higher than the modern. A model run excluding the ecological proxies increases the GMST by `r eoceneproxy_global_diff$"50%"` (`r eoceneproxy_global_diff$"2.5%"` - `r eoceneproxy_global_diff$"97.5%"`)$^\circ$C. The median latitudinal gradient is similar when excluding the ecological proxies, with a median of `r eocene_proxy_gradient$"50%"`$^\circ$C, but with a `r proxy_uncertain_diff_percent`% wider 95% CI (Fig. S2). This indicates that the ecological proxy data are broadly in agreement with the geochemical proxies, while providing additional constraints on the shape of the early Eocene temperature gradient.

Due to the limited spatial coverage of the early Eocene proxy record, and due to the added model complexity of simultaneously estimating a model across both hemispheres, we pooled the proxy data across both hemispheres. Applying the model separately to each hemispheres results in substantial differences in hemispherical, average temperatures, with the southern hemisphere being warmer by `r eeco_south_north_global_diff$"50%"` (`r eeco_south_north_global_diff$"2.5%"` - `r eeco_south_north_global_diff$"97.5%"`)$^\circ$C. The inferred latitudinal gradient is somewhat steeper in the northern hemisphere (steeper by `r eocene_north_south_gradient_diff$"50%"`$^\circ$C, although the 95% CI spans `r eocene_north_south_gradient_diff$"2.5%"` to `r eocene_north_south_gradient_diff$"97.5%"`$^\circ$C), but the large uncertainties associated with both gradients, and the lack of polar proxy data in the southern hemisphere preclude a more precise statement (see Fig. S3).

::: {#fig-4}
![](../figures/fig_4.png){width="10cm"}

Estimates of the median, latitudinal SST gradients of the early Eocene climatic optimum (purple line) and of the present-day (turquoise), both estimated with the Bayesian model. The ribbons (purple and turquoise shadings) depict the 95% credible intervals of these gradients, respectively. Points within the plot depict the geochemical (e.g. &delta;^18^O) and ecological (e.g. mangroves) data. Geochemical data are plotted by their point estimate temperature value. Ecological data are plotted at the mean temperature values of their respective normal distributions.
:::

# Discussion

## Improved estimation of latitudinal and global palaeotemperatures

Our results show that relatively small (*n* = 10 - 20) and patchy samples of temperature data are sufficient to reconstruct greenhouse- and icehouse-type, latitudinal SST gradients, using our Bayesian model. This is an advancement over previously used linear, quadratic or Gaussian approximations [e.g. @Bijl2009; @Tierney2017], which can fit only specific types of gradients. As such, our model presents an alternative to non-parametric methods for inferring latitudinal temperature gradients, which are sometimes favoured as they can very flexibly follow the shape of an unknown temperature gradient [e.g. @Zhang2019; @Jones2022]. However, when used for interpolation or prediction outside the proxy range, non-parametric methods such as Gaussian process regression strictly respond to the data [e.g. @Inglis2020]. This means that the idiosyncrasies of a patchy proxy record, potentially afflicted with measurement errors, calibration errors, and palaeogeographic and temporal uncertainty, directly generates the resulting large-scale climate patterns.

In contrast, our Bayesian, parametric model allows for the inclusion of informative priors on the model parameters. The modeled sea surface temperature gradient thus does not strictly follow the proxy data, but instead represents a compromise between the data and prior knowledge. In the EECO example (Fig. 4), the inclusion of informative priors improves the prediction of sea surface temperatures in the unsampled, very high latitudes: Notice that the upper limit of the credible interval does not increase beyond the range of the data, whereas unconstrained approaches such as splines, Gaussian processes or even standard linear regression could lead to unrealistically high upper bounds in this case [see @Rasmussen2004]. Prior information on the shape of latitudinal temperature gradients on Earth exists for all geological time periods. For example, the greater amount of solar radiation per unit area in low latitudes causes Earth's latitudinal temperature gradient to be broadly negative [@Beer2008]. The ease with which such prior information can be integrated is a major advantage of our method, as the shape of the modeled gradient is controlled by four parameters which clearly relate to its magnitude, steepness and the latitude of its greatest steepness.

Palaeoclimate reconstructions are often summarised as global mean surface temperatures (GMST), providing a standardised metric for characterising the state of the Earth's climate [@Royer2004; @Inglis2020]. The calculation of global mean surface temperatures directly from sparse proxy data is susceptible to bias [@Jones2022]. By modeling the temperature variation across latitudes, a complete temperature distribution along a latitudinal axis can be obtained, filling in gaps in the proxy record through inter- or extrapolation. This eliminates the common problem that specific climate zones dominate the proxy record. Reconstructing the GMST directly from the proxies would lead to an estimate biased towards the well-sampled latitudes. Calculating zonal averages alleviates this problem, but this method relies on comprehensive latitudinal coverage [@Inglis2020]. Instead, our method allows for intersecting the modeled temperature gradient at narrow latitudinal intervals, even when significant latitudinal gaps exist. Weighting the temperatures of those latitudinal intervals by area results in GMST estimates without intrinsic spatial biases. We anticipate that this improved method may significantly alter Phanerozoic, proxy-based temperature curves, which have often been directly calculated from the proxy record [@Royer2004; @Veizer2015]. This is particularly relevant for the early Mesozoic and older intervals, for which the spatial coverage is generally poor due to the absence of data from ocean drilling sites [@Jones2022].

## The role of ecological constraints in palaeoclimate reconstructions

Our results further exemplify how incorporating quantified ecological temperature constraints can provide more precise temperature reconstructions than geochemical proxies alone, adding to the advances in palaeoclimate reconstructions achieved by integrating lithological data [@Scotese2021; @Burgener2023]. Combining the occurrences of climate-sensitive plant communities [@Greenwood1995], reptiles [@Markwick2007], leaf shapes [@Peppe2011], with geochemical proxies offers substantial potential for improving quantitative palaeoclimate reconstructions across the Phanerozoic. Our modeling framework offers a straightforward, efficient way of integrating ecological climate data with other proxy data: The hierarchical model structure accounts for variation of temperature estimates from proxies at individual localities, which is treated equivalent to the uncertainty associated with the ecological temperature proxies. A local temperature estimate, based on multiple geochemical proxies, thus has the same weight as a local temperature estimate obtained from the occurrence of a climate-sensitive plant community, whilst preserving the uncertainty associated with each estimate. The model could easily be extended to include uncertainties on individual geochemical proxy data, or to differently weight proxy records classified as more or less reliable.

Our approach for deriving fully quantitative climate reconstructions from ecological data is borrowed from nearest living relative methods, commonly employed in terrestrial, Cenozoic climate reconstructions [@Fauquette2007; @Pross2012]. One major limitation to these methods is that the thermal preferences of taxa may have changed over time. More significantly, in the early Eocene, sea surface temperatures may have reached heights unknown in the modern world, and nearest living relative methods based on the modern are inherently unable to predict such elevated temperatures. This is especially true for taxa that inhabit the warmest part of the ocean today, e.g. coral reefs [@kleypas1999]. Although coral reefs are threatened by warming sea surface temperatures today [@Hoegh2011], it is conceivable that Eocene reef corals were adapted to a warmer climate. The fossil record indicates that reef development may have been stunted in the early Eocene, with few early Eocene coral reefs occurring in low latitudes [@Zamagni2012]. The absence of coral reefs in higher latitudes in the early Eocene could be due to requirements in irradiance, rather than temperature [@Muir2015]. Tropical temperatures predicted by the geochemical proxy record indicate hotter-than-modern tropical temperatures for the early Eocene (Fig. S2), suggesting that the modern climate range of coral reefs may underestimate the early Eocene thermal niche for coral reefs. We have tried to account for that possibility by widening the temperature probability distribution for reef corals, but the predicted temperatures for the reef and mangrove sites still lie below the temperatures indicated by the geochemical proxy record (Fig. 4, Fig. S2).

<!-- This dilemma could be resolved in future studies by applying machine learning approaches to inferring quantitative temperatures from ecological and lithological proxies: If geochemical proxies and ecological or lithological data are derived from the same location and time interval, the climatic requirements of these taxa or sediments can be learned by using palaeotemperatures from that time interval, instead of or in addition to modern modern temperatures. This approach would ultimately rely on geochemical proxies providing absolute temperature estimates, but the resulting information could be exported to other locations lacking good coverage of geochemical proxies via their fossil or sedimentary record. -->

## Early Eocene climate

The geochemical proxy record and ecological data indicate that the latitudinal SST gradient of the early Eocene climatic optimum was significantly shallower than the modern [@Huber2011], but beyond that, there is little agreement. Earlier, reconstructed early Eocene and EECO SST gradients range from $7 - 21^\circ$C (Table 1); a more recent reconstruction that includes terrestrial air and sea surface temperatures arrives at a gradient of $13^\circ$C [@Inglis2020]. Our polar circle to equatorial gradient estimate is lower than most previous estimates at $7.8^\circ$, although the 95% credible interval extends up to $13.7^\circ$ and thus overlaps earlier estimates based on shallow water proxies. The confirmation of a very flat gradient by both geochemical and ecological shallow water data indicates that inferred SST gradients based on tropical, shallow water and deep water samples [@Cramwinckel2018; @Evans2018] may overestimate the SST gradient of the early Eocene greenhouse world.

<!-- @Evans2018 estimate the SST gradient at 32+-10% of the modern, which would correspond to \~`r round( as.numeric(modern_gradient$"50%") - 0.32*as.numeric(modern_gradient$"50%"),0)`+- `r format(round( 0.10*as.numeric(modern_gradient$"50%"),0),nsmall=0)`$^\circ$C, using the difference between the mean tropical and deep ocean data. The modeled SST of @Tierney2017 using TEX86 is \~ `r round(35.8-23.8,0)`$^\circ$C, when taking the difference between temperatures at the polar circles and the equator. @Inglis2020 include both terrestrial air temperature and SST estimates in their model of the latitudinal temperature gradient in their Figure S4a; they model a polar circle to equatorial gradient of \~ `r round(30.7-(20.2+14.3)/2,0)`$^\circ$C. All of those estimates are significantly lower than the SST gradient predicted by an EECO climate model ensemble [@Lunt2012, as shown in @Tierney2017], which predicts a polar circle to equatorial gradient of `r round(31.5-5.5,0)`$^\circ$C. @Keating-Bitonti2011 show a proxy-based SST gradient of \~ `r round(30-(19+15.6)/2,0)`$^\circ$C, and @Bijl2009 model a gradient of \~`r round(36-27,0)`$^\circ$C in the northern and only \~`r 36-31`$^\circ$C in the southern hemisphere for the early Eocene. Our polar circle to equatorial gradient is lower than most previous estimates at `r eocene_gradient_polar_circle$"50%"`$^\circ$C, although the 95% credible interval overlap with several of the listed gradient estimates. -->

Discrepancies between earlier, proxy-based reconstructions and our modeling results are most pronounced in latitudes beyond the polar circle, as earlier approaches [e.g. @Tierney2017] predict almost linearly decreasing SSTs towards the poles, whereas our median prediction suggests only a slight decrease beyond the polar circle. The scarcity of temperature records in this range leads to widening credible intervals in our prediction, including the possibility of stronger temperature decreases. Polar temperature estimates from our model are thus conservative in that they admit large uncertainty where data is absent, which is desirable. However, the presence of high proxy-derived temperature estimates at \~ 60$^\circ$ latitudes forces the modeled median temperature curve to be too high at \~ 24$^\circ$C, relative to the temperatures indicated by the high-latitude mangrove communities (15.6 - 22.5$^\circ$C). In contrast, the extrapolated polar temperatures of most previous proxy-based models are likely too low, given the abundance of ecological data indicating temperate or subtropical high-latitude climates during the EECO [@Pross2012; @Popescu2021].

The very high variability of the proxy record in mid-latitudes results in large uncertainties on the shape of temperature gradient and on the GMST. Biases and errors in the proxy reconstructions likely contribute to the observed variability, as geochemical proxies reflect many other factors besides seawater temperature [@Hollis2019]. Despite excluding &delta;^18^O measurements from recrystallised fossils, systematic offsets remain between mostly warm temperatures derived from $TEX_{86}$, and cooler temperatures derived from &delta;^18^O, &Delta;~47~, and the ecological proxies. Seasonality [@Keating-Bitonti2011] and temporal changes within the EECO [@Westerhold2018] may also contribute to the large variability of the EECO proxy data.

Recent, marine GMST estimates of the EECO and of the early Eocene range from 23.4 to 37.1$^\circ$C, with the lowest GMSTs being derived from &delta;^18^O, and the higher estimates including $TEX_{86}$ [@Inglis2020]. Many studies include both marine and terrestrial proxies to derive GMST estimates, but despite great differences in proxy selection and in the calculation of global average temperatures, many recent estimates fall in the range of 27 - 29.5$^\circ$C [@Hansen2013; @Caballero2013; @Cramwinckel2018; @Zhu2019], similar to our median GMST estimate of `r eotemp_global$"50%"`$^\circ$C.

# Conclusions

The Bayesian hierarchical model presented herein is able to reconstruct latitudinal gradients from both geochemical and ecological proxy data, whilst reflecting the uncertainty associated with the ecological temperature proxies, and accounts for the variation of multiple temperature estimates at individual localities. Using informative prior information allows for accurate temperature reconstructions from records with geographically incomplete sampling. By providing temperature estimates across the entire latitudinal range, this method also facilitates the reconstruction of unbiased global average temperatures. Application of our model to the EECO suggests that latitudinal sea surface temperature gradients were shallower than estimated by most earlier, proxy-based studies. High-latitude pollen records support this interpretation. Our GMST estimate is in good agreement with most existing estimates, indicating that broadly accurate GMST reconstructions are possible even with substantial deviations in the shape of the latitudinal temperature gradient. Our new method opens the door for improving the accuracy of proxy-based palaeoclimate reoncstructions and Phanerozoic temperature curves, particularly in intervals with a patchy record, and mitigates the biases incurred from uneven sampling. The flexibility of this approach means that our estimates can be quickly and easily updated when new data is made available.

<!-- for reconstructing past temperatures by incorporating both geochemical and ecological data. The use of a Bayesian framework allows for the specification of prior information based on physical constraints, which can improve interpolation and prediction. By modeling the temperature variation across latitudes, a complete temperature distribution is obtained, filling in gaps in the proxy record and reducing the risk of bias in global mean surface temperature (GMST) estimates. The hierarchical structure of the model accounts for the variation of temperature estimates at individual localities and uncertainty associated with the ecological temperature proxies. This improved method is expected to significantly alter the proxy-based temperature reconstructions of the past, particularly for intervals with small or biased samples of temperature proxies. -->

# Acknowledgements

The authors are grateful to all those who have enabled this work by collecting, measuring, collating and screening geochemical and fossil data. The contribution of L.A.J. was supported by a Juan de la Cierva-formación 2021 fellowship (FJC2021-046695-I/MCIN/AEI/10.13039/501100011033) from the European Union "NextGenerationEU"/PRTR. 
<!-- regarding funding acknowledgement, I talked to my boss and he says I don't need to credit my funding as I have done this work largely in my spare time -->

# Author contributions

Both authors designed the study and carried out data preparation. K.E. programmed the model and conducted the analyses. L.A.J. generated the figures. Both authors contributed to the writing of the manuscript.

# Competing Interests

The authors declare that they have no conflicts of interest.

# Data accessibility

The data and code used to produce the results of this study are available via GitHub (<https://github.com/KEichenseer/PalaeoClimateGradient>) and the linked Zenodo repository (XXX).

# References